### 2019-02-06 preliminaries.tex
ssr

\item{(b)} 
Suppose lines $n-1$ and $n$ in the specification of $f$ 
have complementary canalyzed outputs.
Then, the function remains unchanged 
if the canalyzing value and canalyzed output in line $n$
are both complemented. 
(Here, the value on the ``Default" line is also complemented.)
\QED
\end{description}
\end{observation}

@@

\begin{definition}
\label{def:layer}
A {\bf layer} of a NCF representation is a maximal length sequence 
of lines with the same canalyzed output.
\end{definition}
%%Note that the ``Default" line in the representation of an NCF is
%%\emph{not} considered in the definition of a layer.

@@

\medskip
\noindent
This function has 3 layers, with 
Layer~1 consisting of the lines with $x_1$, $x_2$ and $x_3$, 
Layer~2 consisting of the lines with $x_4$ and $x_5$, and
Layer~3 consisting of the line with $x_6$.  \qed

@@

\begin{definition}
\label{def:normalized}
Let $f$ be an NCF with $n \geq 2$ variables specified using
the simplified representation with 
lexicographic ordering given by  
$\langle x_{\pi(1)}, x_{\pi(2)}, \ldots, x_{\pi(n)}\rangle$.
We say that the representation is {\bf normalized} if
both of the following conditions hold.
\begin{enumerate}
  \item Lines $n-1$ and $n$ have the same canalyzed output ~and
  \item each pair of consecutive lines
        with the same canalyzed output occur in the lexicographic order 
        of their canalyzing variables; that is, if two consecutive lines are \\
%\medskip
\noindent
\hspace*{1.1in} $x_i$:~ $a_i \longrightarrow b$ \\
\hspace*{1.1in} $x_j$:~ $a_j \longrightarrow b$

@@

\medskip
\noindent
\textbf{Part 2:}~ Consider the normalized version of Example~4 
produced in Part~1 above. 
In that representation, if we exchange Lines 1 and 2, 
by Part~(a) of Observation~\ref{obs:ncf_transformations},
the function does not change. 
However, the resulting representation is \emph{not} normalized
since the line with $x_2$ precedes the one 
with $x_1$. \qed

\medskip
From Observation \ref{obs:ncf_transformations},
we note that in polynomial time, for any NCF representation with more than one line,
we can first modify the last line if necessary so that it has the same canalyzed output as the preceding line,
and then we can lexicographically sort
each maximal sequence of consecutive lines with the same canalyzed output,
thereby obtaining an equivalent normalized NCF representation.
We state this formally below.

>>>>

\item{(b)} 
Suppose lines $n-1$ and $n$ in the specification of $f$ 
have complementary canalyzed values.
Then, the function remains unchanged 
if the canalyzing value and canalyzed value in line $n$
are both complemented. 
(Here, the value on the ``Default" line is also complemented.)
\QED
\end{description}
\end{observation}

@@

\begin{definition}
\label{def:layer}
A {\bf layer} of a NCF representation is a maximal length sequence 
of lines with the same canalyzed value.
\end{definition}
%%Note that the ``Default" line in the representation of an NCF is
%%\emph{not} considered in the definition of a layer.

@@

\medskip
\noindent
This function has 3 layers, with 
Layer~1 consisting of the lines with $x_1$, $x_2$ and $x_3$, 
Layer~2 consisting of the lines with $x_4$ and $x_5$, and
Layer~3 consisting of the line with $x_6$.  %\qed

@@

\begin{definition}
\label{def:normalized}
Let $f$ be an NCF with $n \geq 2$ variables specified using
the simplified representation with 
lexicographic ordering given by  
$\langle x_{\pi(1)}, x_{\pi(2)}, \ldots, x_{\pi(n)}\rangle$.
We say that the representation is {\bf normalized} if
both of the following conditions hold.
\begin{enumerate}
  \item Lines $n-1$ and $n$ have the same canalyzed value ~and
  \item each pair of consecutive lines
        with the same canalyzed value occur in the lexicographic order 
        of their canalyzing variables; that is, if two consecutive lines are \\
%\medskip
\noindent
\hspace*{1.1in} $x_i$:~ $a_i \longrightarrow b$ \\
\hspace*{1.1in} $x_j$:~ $a_j \longrightarrow b$

@@

\medskip
\noindent
\textbf{Part 2:}~ Consider the normalized version of Example~4 
produced in Part~1 above. 
In that representation, if we exchange Lines 1 and 2, 
by Part~(a) of Observation~\ref{obs:ncf_transformations},
the function does not change. 
However, the resulting representation is \emph{not} normalized
since the line with $x_2$ precedes the one 
with $x_1$. %\qed

\medskip
From Observation \ref{obs:ncf_transformations},
we note that in polynomial time, for any NCF representation with more than one line,
we can first modify the last line if necessary 
so that it has the same canalyzed value as the preceding line,
and then we can lexicographically sort
each maximal sequence of consecutive lines with the same canalyzed value,
thereby obtaining an equivalent normalized NCF representation.
We state this formally below.

### 2019-02-06 symmetry_properties.tex
ssr

In line $k$ of $f$, canalyzing value $a_k$ is either the same or
the complement of canalyzing value $a_j$, and canalyzed output $b_k$
is either the same or the complement of canalyzed output $b_j$,
Thus, there are four possible cases for the form of line $k$.  We
now consider each of the four cases.

@@

Since line $k$ has
the same canalyzing value and canalyzed output as line $j$, line
$k$ must occur in a lower layer than that of line $j$.  Thus, there
is at least one line between line $j$ and line $k$ for which the
canalyzed output is $\overline{b_j}$.  Let line $q$ be the first
such line.  Note that for all $i$ such that $1 \leq i < q$, the
value of variable $x_i$ in assignment $\alpha'$ does not match the
canalyzing value of line $i$, but the value of variable $x_q$ does
match the canalyzing value of line $q$.  Since canalyzed output
$b_q$ equals $\overline{b_j}$, we have that $f(\alpha') =
\overline{b_j}$.

@@

Let line $q$ be the first line such that $j < q \leq k$ and the canalyzed
output of line $q$ is $\overline{b_j}$.  Note that $q$ might possibly
equal $k$.  For all $i$ such that $1 \leq i < q$, the value of
variable $x_i$ in assignment $\alpha'$ does not match the canalyzing
value of line $i$, but the value of variable $x_q$ does match the
canalyzing value of line $q$.  Since canalyzed output $b_q$ equals
$\overline{b_j}$, we have that $f(\alpha') = \overline{b_j}$.

@@

Suppose there is a line below line $j$ for which the canalyzed output is
$\overline{b_j}$.  Let $q$ be the first such line.  Note that for
all $i$ such that $1 \leq i < q$, the value of variable $x_i$ in
assignment $\alpha'$ does not match the canalyzing value of line
$i$, but the value of variable $x_q$ does match the canalyzing value
of line $q$.  Since canalyzed output $b_q$ equals $\overline{b_j}$,
we have that $f(\alpha') = \overline{b_j}$.

@@

Now suppose there is no line below line $j$ for which the canalyzed
output is $\overline{b_j}$.  Then lines $j$ and $k$ both occur in
the last layer of $f$.  Thus, for every line $i$, $1 \leq i \leq
n$, the value of variable $x_i$ in assignment $\alpha'$ does not
match the canalyzing value of line $i$.  Consequently, $f(\alpha')
= \overline{b_j}$, the complement of the canalyzed output in the
last layer of $f$.

@@

Suppose there is a line, other than line $k$, below line $j$ for
which the canalyzed output is $\overline{b_j}$.  Let $q$ be the
first such line.  Note that for all $i$ such that $1 \leq i < q$,
the value of variable $x_i$ in assignment $\alpha'$ does not match
the canalyzing value of line $i$, but the value of variable $x_q$
does match the canalyzing value of line $q$.  Since canalyzed output
$b_q$ equals $\overline{b_j}$, we have that $f(\alpha') =
\overline{b_j}$.

Now suppose that line $k$ is the only line below line $j$ for which
the canalyzed output is $\overline{b_j}$.  Since line $j$ has
canalyzed output $b_j$, line $k$ is the only line in its layer.
Since $f$ is normalized, the last layer of $f$ contains at least
two lines.  Thus, there is a last layer following the layer containing
line $k$.  The canalyzed output of all the lines in this last layer
is $b_j$.  Thus, for every line $i$, $1 \leq i \leq n$, the value
of variable $x_i$ in assignment $\alpha'$ does not match the
canalyzing value of line $i$.  Consequently, $f(\alpha') =
\overline{b_j}$, the complement of the canalyzed output in the last
layer of $f$.  \QED

@@

Let $k' = \pi^{-1}(k)$, so that $\pi(k') = k$.  Canalyzing value
$a_{k'}$ is either the same or the complement of $a_k$, and canalyzed
output $b_{k'}$ is either the same or the complement of $b_k$, Thus,
there are four possible cases for the form of line $k'$ of $f$.  We
will show that in each of the four cases, $f(c_{\pi(1)}, c_{\pi(2)},
\ldots, c_{\pi(n)}) = \overline{b_k}$.

@@

Since line $k'$
has the same canalyzing value and canalyzed output as line $k$, and
$f$ is properly $n$-symmetric, line $k'$ must occur in a lower layer
than that of line $k$.  Thus, there is at least one line between
line $k$ and line $k'$ for which the canalyzed output is $\overline{b_k}$.
Let line $q$ be the first such line.  Note that for all $i$ such
that $1 \leq i < q$, $c_{\pi(i)}$ does not match the canalyzing
value of line $i$, but $c_{\pi(q)}$ does match the canalyzing value
of line $q$.  Since canalyzed output $b_q$ equals $\overline{b_k}$,
we have that $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) =
\overline{b_k}$.

@@

Let line $q$ be the first line such that $k < q \leq k'$ and
the canalyzed output of line $q$ is $\overline{b_k}$.
Note that for all $i$ such that $1 \leq i < q$, 
$c_{\pi(i)}$ does not match the canalyzing value of line $i$,
so $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.

@@

Suppose there is a line below line $k$ for which the canalyzed
output is $\overline{b_k}$.  Let $q$ be the first such line.  Note
that for all $i$ such that $1 \leq i < q$, $c_{\pi(i)}$ does not
match the canalyzing value of line $i$, but $c_{\pi(q)}$ does match
the canalyzing value of line $q$.  Thus, $f(c_{\pi(1)}, c_{\pi(2)},
\ldots, c_{\pi(n)}) = \overline{b_k}$.

Now suppose there is no line below line $k$ for which the canalyzed
output is $\overline{b_k}$.  Then line $k$ occurs in the last layer
of $f$.  Since $f$ is both normalized and properly $n$-symmetric,
this last layer contains exactly two lines.  Thus $k = n-1$ and $k'
= n$.  Note that for all $i$, $c_{\pi(i)}$ does not match the
canalyzing value of line $i$.  Thus, $f(c_{\pi(1)}, c_{\pi(2)},
\ldots, c_{\pi(n)}) = \overline{b_k}$.

@@

Suppose there is a line, other than line $k'$, below line $k$ for
which the canalyzed output is $\overline{b_k}$.  Let $q$ be the
first such line.  Note that for all $i$ such that $1 \leq i < q$,
$c_{\pi(i)}$ does not match the canalyzing value of line $i$, but
$c_{\pi(q)}$ does match the canalyzing value of line $q$.  Thus,
$f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.

Now suppose that line $k'$ is the only line below line $k$ for which
the canalyzed output is $\overline{b_k}$.  Since line $k$ has
canalyzed output $b_k$, line $k'$ is the only line in its layer.
Since $f$ is normalized, there is a last layer following the layer
containing line $k'$.  Thus, for all $i$, $c_{\pi(i)}$ does not
match the canalyzing value of line $i$, so $f(c_{\pi(1)}, c_{\pi(2)},
\ldots, c_{\pi(n)}) = \overline{b_k}$.  \QED

@@

\noindent
Note that if $x_i = 0$ for all $i$, $1 \leq i \leq n$, then the
value of $f_n$ is 0 if $n$ is odd and 1 if $n$ is even.  Also, note
that the above representation is unnormalized. To normalize the
representation, the last line would be changed to have canalyzing
value 0, and the same canalyzed output as the preceding line.

>>>>

In line $k$ of $f$, canalyzing value $a_k$ is either the same or
the complement of canalyzing value $a_j$, and canalyzed value $b_k$
is either the same or the complement of canalyzed value $b_j$,
Thus, there are four possible cases for the form of line $k$.  We
now consider each of the four cases.

@@

Since line $k$ has
the same canalyzing value and canalyzed value as line $j$, line
$k$ must occur in a lower layer than that of line $j$.  Thus, there
is at least one line between line $j$ and line $k$ for which the
canalyzed value is $\overline{b_j}$.  Let line $q$ be the first
such line.  Note that for all $i$ such that $1 \leq i < q$, the
value of variable $x_i$ in assignment $\alpha'$ does not match the
canalyzing value of line $i$, but the value of variable $x_q$ does
match the canalyzing value of line $q$.  Since canalyzed value
$b_q$ equals $\overline{b_j}$, we have that $f(\alpha') =
\overline{b_j}$.

@@

Let line $q$ be the first line such that $j < q \leq k$ and the canalyzed
value of line $q$ is $\overline{b_j}$.  Note that $q$ might possibly
equal $k$.  For all $i$ such that $1 \leq i < q$, the value of
variable $x_i$ in assignment $\alpha'$ does not match the canalyzing
value of line $i$, but the value of variable $x_q$ does match the
canalyzing value of line $q$.  Since canalyzed value $b_q$ equals
$\overline{b_j}$, we have that $f(\alpha') = \overline{b_j}$.

@@

Suppose there is a line below line $j$ for which the canalyzed value is
$\overline{b_j}$.  Let $q$ be the first such line.  Note that for
all $i$ such that $1 \leq i < q$, the value of variable $x_i$ in
assignment $\alpha'$ does not match the canalyzing value of line
$i$, but the value of variable $x_q$ does match the canalyzing value
of line $q$.  Since canalyzed value $b_q$ equals $\overline{b_j}$,
we have that $f(\alpha') = \overline{b_j}$.

@@

Now suppose there is no line below line $j$ for which the canalyzed
value is $\overline{b_j}$.  Then lines $j$ and $k$ both occur in
the last layer of $f$.  Thus, for every line $i$, $1 \leq i \leq
n$, the value of variable $x_i$ in assignment $\alpha'$ does not
match the canalyzing value of line $i$.  Consequently, $f(\alpha')
= \overline{b_j}$, the complement of the canalyzed value in the
last layer of $f$.

@@

Suppose there is a line, other than line $k$, below line $j$ for
which the canalyzed value is $\overline{b_j}$.  Let $q$ be the
first such line.  Note that for all $i$ such that $1 \leq i < q$,
the value of variable $x_i$ in assignment $\alpha'$ does not match
the canalyzing value of line $i$, but the value of variable $x_q$
does match the canalyzing value of line $q$.  Since canalyzed value
$b_q$ equals $\overline{b_j}$, we have that $f(\alpha') =
\overline{b_j}$.

Now suppose that line $k$ is the only line below line $j$ for which
the canalyzed value is $\overline{b_j}$.  Since line $j$ has
canalyzed value $b_j$, line $k$ is the only line in its layer.
Since $f$ is normalized, the last layer of $f$ contains at least
two lines.  Thus, there is a last layer following the layer containing
line $k$.  The canalyzed value of all the lines in this last layer
is $b_j$.  Thus, for every line $i$, $1 \leq i \leq n$, the value
of variable $x_i$ in assignment $\alpha'$ does not match the
canalyzing value of line $i$.  Consequently, $f(\alpha') =
\overline{b_j}$, the complement of the canalyzed value in the last
layer of $f$.  \QED

@@

Let $k' = \pi^{-1}(k)$, so that $\pi(k') = k$.  Canalyzing value
$a_{k'}$ is either the same or the complement of $a_k$, and canalyzed
value $b_{k'}$ is either the same or the complement of $b_k$, Thus,
there are four possible cases for the form of line $k'$ of $f$.  We
will show that in each of the four cases, $f(c_{\pi(1)}, c_{\pi(2)},
\ldots, c_{\pi(n)}) = \overline{b_k}$.

@@

Since line $k'$
has the same canalyzing value and canalyzed value as line $k$, and
$f$ is properly $n$-symmetric, line $k'$ must occur in a lower layer
than that of line $k$.  Thus, there is at least one line between
line $k$ and line $k'$ for which the canalyzed value is $\overline{b_k}$.
Let line $q$ be the first such line.  Note that for all $i$ such
that $1 \leq i < q$, $c_{\pi(i)}$ does not match the canalyzing
value of line $i$, but $c_{\pi(q)}$ does match the canalyzing value
of line $q$.  Since canalyzed value $b_q$ equals $\overline{b_k}$,
we have that $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) =
\overline{b_k}$.

@@

Let line $q$ be the first line such that $k < q \leq k'$ and
the canalyzed value of line $q$ is $\overline{b_k}$.
Note that for all $i$ such that $1 \leq i < q$, 
$c_{\pi(i)}$ does not match the canalyzing value of line $i$,
so $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.

@@

Suppose there is a line below line $k$ for which the canalyzed
value is $\overline{b_k}$.  Let $q$ be the first such line.  Note
that for all $i$ such that $1 \leq i < q$, $c_{\pi(i)}$ does not
match the canalyzing value of line $i$, but $c_{\pi(q)}$ does match
the canalyzing value of line $q$.  Thus, $f(c_{\pi(1)}, c_{\pi(2)},
\ldots, c_{\pi(n)}) = \overline{b_k}$.

Now suppose there is no line below line $k$ for which the canalyzed
value is $\overline{b_k}$.  Then line $k$ occurs in the last layer
of $f$.  Since $f$ is both normalized and properly $n$-symmetric,
this last layer contains exactly two lines.  Thus $k = n-1$ and $k'
= n$.  Note that for all $i$, $c_{\pi(i)}$ does not match the
canalyzing value of line $i$.  Thus, $f(c_{\pi(1)}, c_{\pi(2)},
\ldots, c_{\pi(n)}) = \overline{b_k}$.

@@

Suppose there is a line, other than line $k'$, below line $k$ for
which the canalyzed value is $\overline{b_k}$.  Let $q$ be the
first such line.  Note that for all $i$ such that $1 \leq i < q$,
$c_{\pi(i)}$ does not match the canalyzing value of line $i$, but
$c_{\pi(q)}$ does match the canalyzing value of line $q$.  Thus,
$f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.

Now suppose that line $k'$ is the only line below line $k$ for which
the canalyzed value is $\overline{b_k}$.  Since line $k$ has
canalyzed value $b_k$, line $k'$ is the only line in its layer.
Since $f$ is normalized, there is a last layer following the layer
containing line $k'$.  Thus, for all $i$, $c_{\pi(i)}$ does not
match the canalyzing value of line $i$, so $f(c_{\pi(1)}, c_{\pi(2)},
\ldots, c_{\pi(n)}) = \overline{b_k}$.  \QED

@@

\noindent
Note that if $x_i = 0$ for all $i$, $1 \leq i \leq n$, then the
value of $f_n$ is 0 if $n$ is odd and 1 if $n$ is even.  Also, note
that the above representation is unnormalized. To normalize the
representation, the last line would be changed to have canalyzing
value 0, and the same canalyzed value as the preceding line.
